STJ2012 analysis

### 2016-08-01 analysis/STJ2012_analysis.Rmd
test shore differences for both 97 and 100 OTUs

```{r shorediffs, cache=T}
# Compute within- and between-shore statistics for each Species
shorestats <- data.frame(matrix(dimnames=list(levels(data.frame(sample_data(phy97.f.t))$Species), 
                                              c("Species", "n", "overall", "within", 
                                                "between", "R2", "bd", "p")), ncol=8, nrow=10))
for (i in 1:nlevels(data.frame(sample_data(phy97.f.t))$Species)) {
  sp <- levels(data.frame(sample_data(phy97.f.t))$Species)[i]
  shorestats$Species[i] <- sp
  phy97.f.t.species <- subset_samples(phy97.f.t, Species==sp)  # Set temp phyloseq object
  shorestats$n[i] <- nsamples(phy97.f.t.species)
  md <- meandist(phyloseq::distance(phy97.f.t.species, "bray"), 
                 sample_data(phy97.f.t.species)$Shore)
  shorestats$within[i] <- summary(md)$W  # Weighted mean dissimilarity within both Pools
  shorestats$between[i] <- summary(md)$B  # Mean dissimilarity between Pools
  shorestats$overall[i] <- summary(md)$D  # Overall dissimilarity
  if (nlevels(as(sample_data(phy97.f.t.species), "data.frame")$Shore) > 1) {
    set.seed(70235)
    permanova <- adonis(phyloseq::distance(phy97.f.t.species, "bray") ~ Shore, 
                        data=as(sample_data(phy97.f.t.species), "data.frame"), permutations=999)
    shorestats$R2[i] <- permanova$aov.tab$"R2"[1]  # PERMANOVA partial R-squared
    shorestats$p[i] <- permanova$aov.tab$"Pr(>F)"[1]  # PERMANOVA p-value
    bd <- betadisper(phyloseq::distance(phy97.f.t.species, "bray"),
                     group=as(sample_data(phy97.f.t.species), "data.frame")$Shore)
    shorestats$bd[i] <- TukeyHSD(bd)$group[4]
  } else {
    shorestats$R2[i] <- NA
    shorestats$p[i] <- NA
    shorestats$bd[i] <- NA
  }
}
knitr::kable(shorestats)  # dendrogyra, porites furcata, and siderastrea siderea differ by shore
```

>>>>

```{r shorediffs, cache=T}
shorediffs <- function(phy) {
  # Compute within- and between-shore statistics for each Species
  shorestats <- data.frame(matrix(dimnames=list(levels(data.frame(sample_data(phy))$Species), 
                                                c("Species", "n", "overall", "within", 
                                                  "between", "R2", "bd", "p")), ncol=8, nrow=10))
  for (i in 1:nlevels(data.frame(sample_data(phy))$Species)) {
    sp <<- levels(data.frame(sample_data(phy))$Species)[i]
    shorestats$Species[i] <- sp
    phy.species <- subset_samples(phy, Species==sp)  # Set temp phyloseq object
    shorestats$n[i] <- nsamples(phy.species)
    md <- meandist(phyloseq::distance(phy.species, "bray"), 
                   sample_data(phy.species)$Shore)
    shorestats$within[i] <- summary(md)$W  # Weighted mean dissimilarity within both Pools
    shorestats$between[i] <- summary(md)$B  # Mean dissimilarity between Pools
    shorestats$overall[i] <- summary(md)$D  # Overall dissimilarity
    if (nlevels(as(sample_data(phy.species), "data.frame")$Shore) > 1) {
      set.seed(70235)
      permanova <- adonis(phyloseq::distance(phy.species, "bray") ~ Shore, 
                          data=as(sample_data(phy.species), "data.frame"), permutations=999)
      shorestats$R2[i] <- permanova$aov.tab$"R2"[1]  # PERMANOVA partial R-squared
      shorestats$p[i] <- permanova$aov.tab$"Pr(>F)"[1]  # PERMANOVA p-value
      bd <- betadisper(phyloseq::distance(phy.species, "bray"),
                       group=as(sample_data(phy.species), "data.frame")$Shore)
      shorestats$bd[i] <- TukeyHSD(bd)$group[4]
    } else {
      shorestats$R2[i] <- NA
      shorestats$p[i] <- NA
      shorestats$bd[i] <- NA
    }
  }
  return(shorestats)
}

shorestats97 <- shorediffs(phy97.f.t)
shorestats100 <- shorediffs(phy100.f.t)

knitr::kable(shorestats97, row.names=F, caption="97% OTU data: differences within and between shore for each species")
knitr::kable(shorestats100, row.names=F, caption="100% OTU data: differences within and between shore for each species")
```
